### 2015-04-04 p3_505/paper/paper2.tex

\documentclass[aps,prd,twocolumn,nofootinbib]{revtex4-1}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{fourier}
\usepackage{subfig}
\usepackage{epsfig}
\usepackage{listings}
\usepackage[hidelinks,hyperfootnotes=false,bookmarks=false]{hyperref}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{verbatim}
\usepackage{float}

\begin{document}

\title{Sterile Neutrinos}
\author{Douglas Davis}
\author{Matthew Epland}
\author{Justin Raybern}
\author{Pingchuan Zhao}
\affiliation{Department of Physics, Duke University, Durham, NC 27707, USA}
\date{\today}
\begin{abstract}
an abstract
\end{abstract}\maketitle

\section{theory}
yep

\subsection{a subsection question mark?}
wut

\begin{thebibliography}{9}
\bibitem{ketzer}
  Bernhard Ketzer, \textit{Hybrid Mesons}, Proceedings of Science, Sixth International Conference on Quarks and Nuclear Physics, PoS(QNP2012)025, \url{http://pos.sissa.it/archive/conferences/157/025/QNP2012_025.pdf} %\url{http://arxiv.org/abs/1208.5125}
\bibitem{belle}
  S.~K.~Choi~\emph{et. al.}, \emph{Observation of a Resonancelike Structure in the} $\pi^{+-}\psi^{\prime}$ \emph{Mass Distribution in Exclusive} $B\rightarrow K \pi^{+-}\psi^{\prime}$ \emph{Decays}. Phys. Rev. Lett. \textbf{100}, 142001 8 April 2008.
\bibitem{z4430_dmolecule}
  E.~Braaten~and~M.~Lu, \emph{Line shapes of the Z(4430)} Phys. Rev. D \textbf{79} 051503(R) 11 March 2009
\bibitem{lhcb} % {j4}
  LHCb Collaboration, R. Aaij \textit{et al.}, \textit{Observation of the Resonant Character of the }$Z(4430{)}^{-}$\textit{ State}, Phys. Rev. Lett., \textbf{112}, 22 (2014). \url{http://dx.doi.org/10.1103/PhysRevLett.112.222002}%\url{http://link.aps.org/doi/10.1103/PhysRevLett.112.222002}
\bibitem{j1}
 K.A. Olive \textit{et al.} (Particle Data Group), \textit{hi} (Revised March 2006 by C. Amsler), Chin. Phys. C, 38, 090001 (2014), \url{http://pdg.lbl.gov/2006/reviews/nonqqbar_mxxx050.pdf}
\bibitem{j2}
 Vincent Mathieu, Nikolai Kochelev, and Vicente Vento, \emph{The Physics of Glueballs}, Int. J. Mod. Phys. E \textbf{18}, 1 (2009), \url{http://www.worldscientific.com/doi/abs/10.1142/S0218301309012124}
\bibitem{j3}
 Wolfgang Ochs, \emph{The status of glueballs}, J. Phys. G: Nucl. Part. Phys. \textbf{40}, 043001 (2013), \url{http://dx.doi.org/10.1088/0954-3899/40/4/043001}

\end{thebibliography}

\end{document} %%% end of doc %%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%